How topos theory can help algebra and geometry

### 2018-07-10 slides-como2018.tex
typo

  \jnote{1}{The Kripke--Joyal semantics for the little Zariski topos amounts to
  the following: $\Spec(A) \models \varphi$ iff~$D(1) \models \varphi$, and the
  clauses for~$D(f) \models \varphi$, where~$f$ ranges over the elements
  of~$A$, are given by the following table.
  \[\footnotesize\renewcommand{\arraystretch}{1.25}\begin{array}{@{}l@{\quad}c@{\quad}l@{}}
    D(f) \models \top &\text{iff}& \text{true} \\
    D(f) \models \bot &\text{iff}& \text{$f$ is nilpotent} \\
    D(f) \models x = y &\text{iff}& x = y \in A[f^{-1}] \\
    D(f) \models \varphi \wedge \psi &\text{iff}&
      \text{$D(f) \models \varphi$ and $D(f) \models \psi$} \\
    D(f) \models \varphi \vee \psi &\text{iff}&
      \text{there exists a partition~$f^n = fg_1 + \cdots + fg_m$ with,} \\
    &&\quad\text{for each~$i$, $D(fg_i) \models \varphi$ or $D(fg_i) \models \psi$} \\
    D(f) \models \varphi \Rightarrow \psi &\text{iff}&
      \text{for all~$g \in A$, $D(fg) \models \varphi$ implies $D(fg) \models \psi$} \\
    D(f) \models \forall x\?M^\sim\_ \varphi(x) &\text{iff}&
      \text{for all~$g \in A$ and $x_0 \in M[(fg)^{-1}]$, $D(fg) \models \varphi[x_0/x]$} \\
    D(f) \models \exists x\?M^\sim\_ \varphi(x) &\text{iff}&
      \text{there exists a partition~$f^n = fg_1 + \cdots + fg_m$ with,} \\
    &&\quad\text{for each~$i$, $D(fg_i) \models \varphi[x_0/x]$ for some~$x_0 \in M[(fg_i)^{-1}]$}
  \end{array} \]
  The generic prime filter~$\fff_0$ can also be described in explicit
  terms. For ring elements~$f$ and~$s$, $D(f) \models (s \in \fff_0)$ iff~$f
  \in \sqrt{(s)}$.}

>>>>

  \jnote{1}{The Kripke--Joyal semantics for the little Zariski topos amounts to
  the following: $\Spec(A) \models \varphi$ iff~$D(1) \models \varphi$, and the
  clauses for~$D(f) \models \varphi$, where~$f$ ranges over the elements
  of~$A$, are given by the following table.
  \[\footnotesize\renewcommand{\arraystretch}{1.25}\begin{array}{@{}l@{\quad}c@{\quad}l@{}}
    D(f) \models \top &\text{iff}& \text{true} \\
    D(f) \models \bot &\text{iff}& \text{$f$ is nilpotent} \\
    D(f) \models x = y &\text{iff}& x = y \in A[f^{-1}] \\
    D(f) \models \varphi \wedge \psi &\text{iff}&
      \text{$D(f) \models \varphi$ and $D(f) \models \psi$} \\
    D(f) \models \varphi \vee \psi &\text{iff}&
      \text{there exists a partition~$f^n = fg_1 + \cdots + fg_m$ with,} \\
    &&\quad\text{for each~$i$, $D(fg_i) \models \varphi$ or $D(fg_i) \models \psi$} \\
    D(f) \models \varphi \Rightarrow \psi &\text{iff}&
      \text{for all~$g \in A$, $D(fg) \models \varphi$ implies $D(fg) \models \psi$} \\
    D(f) \models \forall x\?M^\sim\_ \varphi(x) &\text{iff}&
      \text{for all~$g \in A$ and $x_0 \in M[(fg)^{-1}]$, $D(fg) \models \varphi(x_0)$} \\
    D(f) \models \exists x\?M^\sim\_ \varphi(x) &\text{iff}&
      \text{there exists a partition~$f^n = fg_1 + \cdots + fg_m$ with,} \\
    &&\quad\text{for each~$i$, $D(fg_i) \models \varphi(x_0)$ for some~$x_0 \in M[(fg_i)^{-1}]$}
  \end{array} \]
  The generic prime filter~$\fff_0$ can also be described in explicit
  terms. For ring elements~$f$ and~$s$, $D(f) \models (s \in \fff_0)$ iff~$f
  \in \sqrt{(s)}$.}

### 2018-07-10 slides-leipzig2018.tex
typo

  \vspace*{-1em}
  \[\footnotesize\renewcommand{\arraystretch}{1.25}\begin{array}{@{}l@{\ }c@{\ }l@{}}
    D(f) \models \top &\text{iff}& \text{true} \\
    D(f) \models \bot &\text{iff}& \text{$f$ is nilpotent} \\
    D(f) \models x = y &\text{iff}& x = y \in A[f^{-1}] \\
    D(f) \models \varphi \wedge \psi &\text{iff}&
      \text{$D(f) \models \varphi$ and $D(f) \models \psi$} \\
    D(f) \models \varphi \vee \psi &\text{iff}&
      \text{there exists a partition~$f^n = fg_1 + \cdots + fg_m$ with,} \\
    &&\quad\text{for each~$i$, $D(fg_i) \models \varphi$ or $D(fg_i) \models \psi$} \\
    D(f) \models \varphi \Rightarrow \psi &\text{iff}&
      \text{for all~$g \in A$, $D(fg) \models \varphi$ implies $D(fg) \models \psi$} \\
    D(f) \models \forall x\?M^\sim\_ \varphi(x) &\text{iff}&
      \text{for all~$g \in A$ and $x_0 \in M[(fg)^{-1}]$, $D(fg) \models \varphi[x_0/x]$} \\
    D(f) \models \exists x\?M^\sim\_ \varphi(x) &\text{iff}&
      \text{there exists a partition~$f^n = fg_1 + \cdots + fg_m$ with,} \\
    &&\quad\text{for each~$i$, $D(fg_i) \models \varphi[x_0/x]$ for some~$x_0 \in M[(fg_i)^{-1}]$}
  \end{array} \]
\end{frame}

@@

>>>>

  \vspace*{-1em}
  \[\footnotesize\renewcommand{\arraystretch}{1.25}\begin{array}{@{}l@{\ }c@{\ }l@{}}
    D(f) \models \top &\text{iff}& \text{true} \\
    D(f) \models \bot &\text{iff}& \text{$f$ is nilpotent} \\
    D(f) \models x = y &\text{iff}& x = y \in A[f^{-1}] \\
    D(f) \models \varphi \wedge \psi &\text{iff}&
      \text{$D(f) \models \varphi$ and $D(f) \models \psi$} \\
    D(f) \models \varphi \vee \psi &\text{iff}&
      \text{there exists a partition~$f^n = fg_1 + \cdots + fg_m$ with,} \\
    &&\quad\text{for each~$i$, $D(fg_i) \models \varphi$ or $D(fg_i) \models \psi$} \\
    D(f) \models \varphi \Rightarrow \psi &\text{iff}&
      \text{for all~$g \in A$, $D(fg) \models \varphi$ implies $D(fg) \models \psi$} \\
    D(f) \models \forall x\?M^\sim\_ \varphi(x) &\text{iff}&
      \text{for all~$g \in A$ and $x_0 \in M[(fg)^{-1}]$, $D(fg) \models \varphi(x_0)$} \\
    D(f) \models \exists x\?M^\sim\_ \varphi(x) &\text{iff}&
      \text{there exists a partition~$f^n = fg_1 + \cdots + fg_m$ with,} \\
    &&\quad\text{for each~$i$, $D(fg_i) \models \varphi(x_0)$ for some~$x_0 \in M[(fg_i)^{-1}]$}
  \end{array} \]
\end{frame}

\backupstart

@@

\backupend
